---
title: "Statistical Analysis of COVID-19"
author: "Ayorinde Ayomide David"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
if(!require(pacman)) install.packages("pacman")

pacman::p_load(tidyverse, # Metapackage
               here, # Used to coerce Rmarkdown into reading our dataset from a seperate folder
               Hmisc
               )
```

### Loading the COVID-19 dataset into R

```{r Data, message=FALSE, warning=FALSE}

covid <- read_csv(here("Data/COVID19_line_list_data.csv"))
```

### Exploring and inspecting the dataset

```{r message=FALSE, warning=FALSE}
view(covid)

describe(covid)

covid %>% 
  select(death) %>% 
  unique()
```

We have 14 unique entries for death, however, death should have 2 unique values namely:

-   0 - Alive
-   1 - Dead

Hence, the need to clean the column to clear out the inconsistencies

Cleaning the column of interest

```{r}
names(covid)
covid <- covid %>% 
  select(!c(...4,...22,...23,...24,...25,...26)) %>%  #removing columns with all observation missing
  
  mutate(death_cleaned = as.integer(covid$death != 0))

```

#### Checking to confirm if the cleaning has been effected

```{r}
covid %>% 
  select(death_cleaned) %>% 
  unique()

```

#### Calculating death rate

```{r Death rate}
covid %>% 
  select(death_cleaned) %>% 
  sum()/ nrow(covid)

```

**Statistical Analysis: Two-sample t-test**

H~0~: There is no significant difference between the age of those alive and dead ($\mu_1 = \mu_2$)

H~1~: There is a significant difference between the age of those alive and dead ($\mu_1 \ne \mu_2$)

```{r}
covid %>% 
  group_by(death_cleaned) %>% 
  summarise(gender_death_mean = mean(age, na.rm = T))

```

We can see that there is a difference of about 20(in years) between the ages of those that are dead and alive. Now, the question:

Is this really significant?

Let's confirm using **t.test** function

```{r}
dead <- covid %>% 
  filter(death_cleaned == 1)
alive <- covid %>% 
  filter(death_cleaned == 0)

t.test(alive$age,
       dead$age,
       conf.level = 0.95,
       alternative = "two.sided")


```

**Decision rule**: If p-value is \< 0.05, we reject null hypothesis, otherwise, we fail to reject null hypothesis

**Conclusion**: Since the p-value is \< 0.05, we reject null hypothesis and conclude that there is a significant difference between the age of those that are dead and those that are alive. In order words, older people are more likely/prone to death if tested positive for COVID-19

**Test of Independence**

```{r}
covid <- covid %>% 
  drop_na(gender) %>% 
  mutate(gender_cov = factor(gender),
         death_cov = factor(case_when(death_cleaned == 1 ~ "Alive",
                               death_cleaned == 0 ~ "Dead")))

ggplot(covid, mapping = aes(x = gender_cov,
                            fill = death_cov)) +
  geom_bar() +
  labs(title = "Chi-Square Test of Independence Bar Chart",
       subtitle = "Gender and Death",
       x = "Gender",
       caption = "Author: Ayorinde Ayomide David",
       fill = "Life status")

```




Are the proportions of gender independent of life status?

The question above leads us to the hypothesis below

H~0~: The variables are independent i.e There is no relationship between the variables

H~1~: The variables are not independent i.e There is a relationship between the variables

```{r}
covid %>% 
  select(gender_cov,
         death_cov) %>% 
  table() %>% 
  chisq.test()

  
```

**Decision rule**: If p-value is \< 0.05, we reject null hypothesis, otherwise, we fail to reject null hypothesis

**Conclusion**: Since the p-value is \< 0.05, we reject null hypothesis and conclude that there is a relationship between the death and gender variable

```{r}

mutate
```

```{r}
```
